---
title: "Homework 4"
author: "David"
date: "`r Sys.time()`"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---
<!---
In order to check that all the necessary packages are installed, Click the File > New File > R Markdown button and keep all of the options as is. Save the document somewhere, and, while that new document is open, click the knit button. If a window with some plots shows up, everything is installed correctly. If RStudio prompts you to install any packages, you should install them.
-->

<!---
Please save this file as Homework4_lastname_firstname.Rmd

Be sure to update the author field in the block above.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)  
library(car)
library(ggpubr)
library(psych)
library(knitr)

set.seed(1) 
theme_set(theme_minimal())

```

# Part 0: Importing Data  

Import the “participant_level_data_HW4.csv” dataframe supplied on Canvas.
```{r}
participant_data <- read.csv(file.path("data", "participant_level_data_HW4.csv"))
```


   
# Part 1: Correlations 
  
## Question 1:  

Calculate a correlation between “true_recall_proportion” and “false_recall_proportion” from scratch, save the correlation coefficient as “corr_coef” and output this value.  

```{r}

# Pearson is just cov(x,y) / (sd_x * sd_y)

# Get the data
true_recall_values <- participant_data$true_recall_proportion
false_recall_values <- participant_data$false_recall_proportion

# Get means
true_mean <- mean(true_recall_values)
false_mean <- mean(false_recall_values)

# Center data around mean to make cov calculatin easier
true_centered <- true_recall_values - true_mean
false_centered <- false_recall_values - false_mean

# Get standard devs
true_sd <- sd(true_recall_values)
false_sd <- sd(false_recall_values)

# Get covariance
covariance <- sum(true_centered * false_centered) / (length(true_recall_values) - 1)

# Get correlation coefficient
corr_coeff <- covariance / (true_sd * false_sd)

# Log the results
cat(sprintf("Correlation Coefficient (True Recall & False Recall): %.6f", corr_coeff))



```

  
## Question 2:  

Let’s see how our continuous variables relate to each other. Create and show a correlation matrix comparing the following values. You can use built-in functions to calculate correlations.  

a. True recall proportion  
b. False recall proportion  
c. VVIQ score  

```{r}

# Get all the columns of interest
relevant_columns <- participant_data[, c("true_recall_proportion",
                                         "false_recall_proportion",
                                         "VVIQ_score")]

# Get the correlation matrix
correlation_matrix <- cor(relevant_columns)

# Log the results
knitr::kable(
  correlation_matrix,
  caption = "Correlation Matrix for True Recall Proportion, False Recall Proportion, and VVIQ Score"
)

```

*Briefly discuss: Describe the significant relationships you see in this matrix*
  The diagonal (top left to bottom right) is always exactly 1 because every variable relates perfectly to itself. The matrix is also symmetric across the diagonal as it shows each relationship twice. True recall has a 0.14 relationship with false recall, and this is the strongest correlation between these three variables. VVIQ and true recall have a near zero correlation. VVIQ and false recall share a negative relationship of a similar magnitude as true recall and false recall.



# Part II: Cross Validation  
  
## Question 3:  

Perform a hold-out cross-validation of the mixed effects regression model you made in HW 3, Part II, Question 13. 

### 3.a:  
Set your seed to 2026. Split your data randomly into 80% training / 20% testing. Output the first 10 rows of the test and training datasets.   
```{r}

# Load libs
library(lme4)
library(lmerTest)

# Load class data
class_data <- read.csv(file.path("data", "class_data.csv"))

# Standardize variables
class_data$true_recall_proportion_z <- as.numeric(scale(class_data$true_recall_proportion))
class_data$vviq_z <- as.numeric(scale(class_data$VVIQ_score))

# Convert predictors to factors
class_data$encoding_strategy <- as.factor(class_data$encoding_strategy)
class_data$interference_type <- as.factor(class_data$interference_type)
class_data$participant <- as.factor(class_data$participant)

# Get unique participant IDs
participant_ids <- unique(class_data$participant)
num_participants <- length(participant_ids)

# Set the seed
set.seed(2026)

# Number of splits we will do and test/train fractions
num_splits <- 1000
train_fraction <- 0.80
test_fraction <- 0.20

# Get number of participants in training set
num_train_participants <- floor(train_fraction * num_participants)

# Randomly select training participants (we gonna use it as a mask)
train_participants <- sample(participant_ids,
                             size = num_train_participants,
                             replace = FALSE)

# Create training and testing datasets
training_data <- class_data[class_data$participant %in% train_participants, ]
testing_data  <- class_data[!(class_data$participant %in% train_participants), ]

# Output first 10 rows
head(training_data, 10)
head(testing_data, 10)

```
  
### 3.b:  
Use the training data to determine the best fit model and output the summary.
```{r}

# Mixed effects model from HW3
mixed_model_training <- lmer(
  true_recall_proportion_z ~
    encoding_strategy +
    interference_type +
    vviq_z +
    (1 | participant),
  data = training_data
)

# Print da results
summary(mixed_model_training)
```
  
### 3.c:  
Calculate the model accuracy with the testing data (to measure performance, let’s use r, the correlation between the predictions and true values).   
```{r}

# Get the predicted values
predicted_test_values <- predict(
  mixed_model_training,
  newdata = testing_data,
  allow.new.levels = TRUE
)

# Calculate correlation between predicted and actual values
test_r <- cor(predicted_test_values,
              testing_data$true_recall_proportion_z)

cat(sprintf("Test-set correlation (r): %.6f\n", test_r))
```

  
### 3.d:  
Conduct this cross-validation over 1000 splits. Call the list of resulting model accuracy values cv_r. Set the seed to 2026 before starting your 1000 splits   
```{r}

# Set the seed
set.seed(2026)

# Intialize empty list of proper size
cv_r <- rep(NA, num_splits)

# Iterate number of time we do the split
for (split_index in 1:num_splits) {

  # Get number of training participants
  num_train_participants <- floor(train_fraction * num_participants)
  
  # Get the ids for the participants that will be in training set
  train_participants <- sample(participant_ids,
                               size = num_train_participants,
                               replace = FALSE)
  
  # Get training and testing data
  training_data <- class_data[class_data$participant %in% train_participants, ]
  testing_data  <- class_data[!(class_data$participant %in% train_participants), ]
  
  # Train that sucker 
  mixed_model_training <- lmer(
    true_recall_proportion_z ~
      encoding_strategy +
      interference_type +
      vviq_z +
      (1 | participant),
    data = training_data,
  )
  
  # Predict them values
  predicted_test_values <- predict(
    mixed_model_training,
    newdata = testing_data,
    allow.new.levels = TRUE
  )

  # Calculate correlation between predicted and actual values
  test_r <- cor(predicted_test_values, testing_data$true_recall_proportion_z)
  
  # Insert the correlation coefficient in the proper index
  cv_r[split_index] <- test_r
}
```

  
### 3.e:  
Plot a histogram of your model accuracy values and report the mean model accuracy. Also save the mean model accuracy to a variable called `mean_cv_r`. 
```{r}

# Get a little histogram action goin
hist(cv_r,
     main = "Cross-Validation Accuracy (80/20)",
     xlab = "Model Accuracy (r)")

# Get mean of the cross validation r values
mean_cv_r <- mean(cv_r)

# Print it
cat(sprintf("Mean Cross-Validation Accuracy (80/20): %.6f\n", mean_cv_r))
```

  
### 3.f:  
Transform the r values with a Fisher Z-transformation (call this cv_z) and conduct a Nadeau-Bengio-corrected t-test of the values versus 0. Save the results in a variable called `cv_t`.
```{r}

bengio.t.test <- function(results, nperm, ntrain, ntest, alternative="two.sided") {
  
  m = mean(results)
  var_res = sd(results)**2
  adjusted_var = ((1/nperm) + (ntest/ntrain))*var_res
  
  t = m / sqrt(adjusted_var)
  
  df = nperm - 1
  
  # There are lots of ways to get to p-values from here
  if (alternative=="less") {
    pval = pt(t, df=df) 
  } else if (alternative=="greater") {
    pval = 1 - pval
  } else if (alternative=="two.sided") {
    pval = pt(abs(t), df=df, lower.tail=F) * 2
  }
  
  return(data.frame(t=t, pval=pval, df=df))
}

```

```{r}

# Write your code here

```
  
### 3.g:  
*Briefly discuss: Interpret the resulting t-test and p-value. What does this say about your model?*   
  
  
 
  
## Question 4:  
Conduct a hold-out cross-validation of your model the same as above, but with 50% training and 50% testing. Set the seed to 2026 before starting and then save the model accuracy results in a variable called `cv_r_half`.  

```{r}

set.seed(2026)

```

### 4.a:
*Briefly discuss: How does the ability of your model to extend to other data change? What does this say about the ideal distribution of training and testing data?*  
  

# Part III: ANOVAs  
  
## Question 5:  
This study was designed to be analyzed with an ANOVA (it’s originally designed for a 3-way ANOVA, but let’s start simpler with a 2-way ANOVA). In particular, let’s see how VVIQ score (imagery ability) interacts with encoding depth to impact memory performance. First, we need to make a categorical version of the VVIQ. Create a new column in the dataframe called vviq_level where it is 1 if the participant’s VVIQ score is equal to or higher than the median score across participants, and 0 if it is lower than the median score across participants.  

Calculate a 2-way independent samples ANOVA with the DV as true recall performance, and the IVs varying along 2 factors: 1) vviq (categorical, as low / high) and 2) encoding strategy depth (categorical, as deep / shallow).

### 5.a:  
Calculate Sum of Squares for vviq from scratch. Call this *SS_vviq* and output this value.  
```{r}

```

  
### 5.b:  
Calculate Sum of Squares for strategy from scratch. Call this *SS_strategy* and output this value.  
```{r}

```

  
### 5.c:  
Calculate Sum of Squares total from scratch. Call this *SS_total* and output this value.  
```{r}

```

  
### 5.d:  
Calculate Sum of Squares within from scratch. Call this *SS_within* and output this value.  
```{r}

```



## Question 6:  
Write code that generates an ANOVA table that uses the format from the HW 4 instructions document, and populate it using your calculations. Calculate MS and F from scratch. The p-values can be calculated with built-in R functions. Output this table.  
(hint: feel free to use knitr::kable)

**For the five cells that have a “--” in the HW instructions, leave these in your dataframe as “NA”s.**  

```{r}
```


  
## Question 7:  
Create a bar graph showing the means of the 4 different conditions (2 VVIQ levels x 2 strategy types), and mark significance for the factors/interactions on this plot. 
```{r}

```

  
  
## Question 8:  
Calculate effect size (partial eta squared) for each factor and the interaction from scratch. Call these: `eta_vviq_level`, `eta_strategy`, and `eta_interaction` and output these values.  
```{r}

```
  
  
  
## Question 9:  
Describe the statistics for each factor, their interaction, and the effect sizes as if for a research paper, referring to this bar graph. Include interpretation of the results.  

*Please write your response in the lines below.*  
  

  
## Question 10:  
Calculate this ANOVA using R built-in functions to check your numbers.  
```{r}

```
  


## Question 11:  
Calculate this ANOVA instead as a within-subjects (repeated measures) ANOVA using R built-in functions. Save the results in a variable called anova_within and display the statistic and p values for each factor and their interaction.  
```{r}

```

*Briefly discuss: How do the results change when instead using a repeated measures ANOVA? What is the correct function to use?*  
  
  

# Part IV: ANOVA as regression  


## Question 12:
Write the independent samples ANOVA as a regression model, and report the regression model formula. (Remember that VVIQ and encoding strategy are now categorical, not continuous.)  
```{r}

```



## Question 13:  
Conduct the regression model using R built-in functions. How do the outputs of this model compare to those of the ANOVA?
```{r}
```
